---
title: "cleanYourFrame"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cleanYourFrame}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r message=F, warning=F}
library(frameCleaneR)
library(printr)
library(dplyr)
```

Data imported from excel and csv in business situations can have messy characteristics and data formats. This package provides functions to tidy your data frame using the power of `tidyselect`.

**create sample data**

```{r}
tibble::tibble(
  date = c("20190101", "20190305", "20201012"),
  numeric_val = c(1, NA, 5),
  char_val = c("", "    val ", "-")
) -> sample_table

sample_table
```

# set nas

Data occasionally has different ways to represent NA values. `set_na` checks as default `c("-", "", " ", "null")` but any values can be supplied to automatically be set to NA. This is helpful when you want to check the NA profile of a data frame using `dlookr::diagnose`

```{r}
sample_table %>% 
  set_na(everything())
```

# remove whitespace

remove whitespace from the ends of character variables that may be otherwise undetectable by inspection.

```{r}
sample_table %>% 
  remove_whitespace(where(is.character))
```

# set dates

automatically convert character columns that should be dates.

```{r}
sample_table %>% 
  set_dates(everything())
```
# relocate all

relocates an unorganized dataframe using heuristics such as putting character and date columns first, and organizing by alphabetical order. 

```{r}
sample_table %>% 
  relocate_all()
```


# clean frame

Wrapper function to apply all cleaning operations to a data frame using sensible defaults. I encourage you to build your own `clean_frame` function to suit your needs. 

```{r}
sample_table %>% 
  clean_frame()
```

# fill nas

use tidyselect to fill NAs with a single value

```{r}
sample_table %>% 
  fill_na(where(is.numeric))
```
# Pivot summary

Transpose a tibble of summary statistics in tidy format. Convenient function for transposing the output of dplyr"s `group_by` and `summarize` operation.

## 0 groups

Transpose a 1 row numerical summary:

**wide format**

```{r}
iris %>% 
  summarize(across(where(is.numeric), mean), .groups = "drop") -> sumr0

sumr0
```

**long format**

```{r}

sumr0 %>% 
  pivot_summary()

```

## 1 group

A grouped summary can be transposed by providing the name of the group column.

**long format**

```{r}
iris %>% 
  group_by(Species) %>% 
  summarize(across(where(is.numeric), mean), .groups = "drop") -> sumr1

sumr1
```
**wide format**

```{r}

sumr1 %>% 
  pivot_summary(Species)
```


## 2 groups

Supports transposing numerical summaries with multiple groups using tidyselect. 

**long format**


```{r}

iris %>%
  mutate(Species1 = sample(Species)) %>%
  group_by(Species, Species1) %>% 
  summarize(across(where(is.numeric), mean), .groups = "drop") -> sumr2
 
sumr2
```


Group names are concatenated and pivoted.

**wide format**

```{r}
sumr2 %>% 
   pivot_summary(matches("Spec")) 
```

